\documentclass[a4paper, 11pt]{article}
\usepackage{comment} % enables the use of multi-line comments (\ifx \fi) 
\usepackage{xparse}% http://ctan.org/pkg/xparse
\NewDocumentCommand{\Log}{o}{%
  \IfNoValueTF{#1}{}{{}^{#1}\!}\log}%
\usepackage{fullpage} % changes the margin
\usepackage{longtable}
\usepackage{graphicx}
\usepackage{fancyvrb,xcolor}
\usepackage{listings}
\usepackage{color}
\usepackage[hyphenbreaks]{breakurl}
\usepackage[hyphens]{url}
\usepackage[margin=3cm]{geometry}
\usepackage{relsize}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\usepackage{float}
\usepackage{caption}
\usepackage{url}
\usepackage{biblatex}
\DeclareCaptionFont{white}{\color{white}}
\DeclareCaptionFormat{listing}{\colorbox{gray}{\parbox{\textwidth}{#1#2#3}}}
\captionsetup[lstlisting]{format=listing,labelfont=white,textfont=white}
\newcommand{\bigqm}[1][1]{\text{\larger[#1]{\textbf{?}}}}
\lstset{
  language=Java,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3
}
\graphicspath{ {images/} }

\begin{document}
%Header-Make sure you update this information!!!!
\noindent
\large\textbf{Assignment 7} \hfill \textbf{Hussam Hallak} \\
\normalsize CS532, Web Science, Spring 2017\hfill CS Master's Student \\
Old Dominion University, Computer Science Dept \hfill Prof: Dr. Nelson 

\section*{Question 1:}
The goal of this project is to use the basic recommendation principles
we have learned for user-collected data. You will modify the code
given to you which performs movie recommendations from the MovieLense
data sets.

The MovieLense data sets were collected by the GroupLens Research
Project at the University of Minnesota during the seven-month period
from September 19th, 1997 through April 22nd, 1998.  We are using the 
"100k dataset"; available for download from:

http://grouplens.org/datasets/movielens/100k/

There are three files which we will use:

1.  u.data: 100,000 ratings by 943 users on 1,682 movies. Each
user has rated at least 20 movies. Users and items are numbered
consecutively from 1. The data is randomly ordered. This is a tab
separated list of 

user id | item id | rating | timestamp

The time stamps are unix seconds since 1/1/1970 UTC.

Example:

196     242     3       881250949
186     302     3       891717742
22      377     1       878887116
244     51      2       880606923
166     346     1       886397596
298     474     4       884182806
115     265     2       881171488

2.  u.item: Information about the 1,682 movies. This is a tab
separated list of

movie id | movie title | release date | video release date | IMDb URL | unknown | Action | Adventure | Animation |Children's | Comedy | Crime | Documentary | Drama | Fantasy | Film-Noir | Horror | Musical | Mystery | Romance | Sci-Fi | Thriller | War | Western |

The last 19 fields are the genres, a 1 indicates the movie is of
that genre, a 0 indicates it is not; movies can be in several genres
at once. The movie ids are the ones used in the u.data data set.

Example:

161|Top Gun (1986)|01-Jan-1986||http://us.imdb.com/M/title-exact?Top%20Gun%20(1986)|0|1|0|0|0|0|0|0|0|0|0|0|0|0|1|0|0|0|0 
162|On Golden Pond (1981)|01-Jan-1981||http://us.imdb.com/M/title-exact?On%20Golden%20Pond%20(1981)|0|0|0|0|0|0|0|0|1|0|0|0|0|0|0|0|0|0|0 
163|Return of the Pink Panther, The (1974)|01-Jan-1974||http://us.imdb.com/M/title-exact?Return%20of%20the%20Pink%20Panther,%20The%20(1974)|0|0|0|0|0|1|0|0|0|0|0|0|0|0| 0|0|0|0|0

3.  u.user: Demographic information about the users. This is a tab
separated list of:

user id | age | gender | occupation | zip code

The user ids are the ones used in the u.data data set.

Example:

1$|$24$|$M$|$technician$|$85711 

2$|$53$|$F$|$other$|$94043 

3$|$23$|$M$|$writer$|$32067 

4$|$24$|$M$|$technician$|$43537 

5$|$33$|$F$|$other$|$15213

The code for reading from the u.data and u.item files and creating
recommendations is described in the book Programming Collective
Intelligence.  Feel free to modify the PCI code to answer the 
following questions.

Questions (10 points).

1.  Find 3 users who are closest to you in terms of age, 
gender, and occupation.  For each of those 3 users:

- what are their top 3 favorite films?
- bottom 3 least favorite films?

Based on the movie values in those 6 tables (3 users X (favorite +
least)), choose a user that you feel is most like you.  Feel 
free to note any outliers (e.g., "I mostly identify with user 123,
except I did not like ``Ghost'' at all").  

This user is the "substitute you".  

2.  Which 5 users are most correlated to the substitute you? Which
5 users are least correlated (i.e., negative correlation)?

3.  Compute ratings for all the films that the substitute you
have not seen.  Provide a list of the top 5 recommendations for films
that the substitute you should see.  Provide a list of the bottom
5 recommendations (i.e., films the substitute you is almost certain
to hate).

4.  Choose your (the real you, not the substitute you) favorite and
least favorite film from the data.  For each film, generate a list
of the top 5 most correlated and bottom 5 least correlated films.
Based on your knowledge of the resulting films, do you agree with
the results?  In other words, do you personally like / dislike
the resulting films?

\subsection*{Answer:}
1. In order to find 3 users who are closest to me in terms of age, 36 years old, and gender, male, and occupation, programmer. I wrote a small python script and named it ``findclosestusers.py''. I kept refining the results until I got the 3 closest users. The script takes a text file as input from the command line and saves the results to a text file named ``closest3users.txt''.

\lstinputlisting[language=Python, breakatwhitespace=〈false), label=The content of findclosestusers.py, caption= The content of findclosestusers.py]{Q1/findclosestusers.py}

After running the script ``findclosestusers.py'' and passing the file ``u.user'', I got the following users:

\lstinputlisting[language=Python, breakatwhitespace=〈false), label=The content of closest3users.txt, caption= The content of closest3users.txt]{Q1/closest3users.txt}


In order to find the top 3 and the bottom 3 favorite films for each of the 3 users, I wrote a python script, ``findtopbottom343.py''. The program takes 3 command line arguments:

a. The file ``closest3users.txt'': Gives users' ids.

b. The file ``u.data'': Gives items' ids and ratings that belong to the 3 users.

c. The file ``u.item'' Gives the data for the 6 films' we need.

The output is saved to a file named ``topbottom343.txt''

\lstinputlisting[language=Python, breakatwhitespace=〈false), label=The content of findtopbottom343.py, caption= The content of findtopbottom343.py]{Q1/findtopbottom343.py}

After running the script, I got the following output:

\begin{lstlisting}[language=bash, breakatwhitespace=〈false), label=Running findtopbottom343.py, caption=Running findtopbottom343.py]
root@ima-app:/var/www/Hussam/A7# python findtopbottom343.py closest3users.txt u.data u.item
root@ima-app:/var/www/Hussam/A7# cat topbottom343.txt
First User id: 182
******************
Top 3 films:
--------------
Star Wars (1977)
Hunchback of Notre Dame, The (1996)
12 Angry Men (1957)
--------------------------------------------
Bottom 3 films:
-----------------
Emma (1996)
That Thing You Do! (1996)
Donnie Brasco (1997)
*****************************************************
Second User id: 215
******************
Top 3 films:
--------------
Silence of the Lambs, The (1991)
Wizard of Oz, The (1939)
Searching for Bobby Fischer (1993)
--------------------------------------------
Bottom 3 films:
-----------------
Unbearable Lightness of Being, The (1988)
Seven (Se7en) (1995)
Star Trek V: The Final Frontier (1989)
*****************************************************
Third User id: 633
******************
Top 3 films:
--------------
Die Hard 2 (1990)
Basic Instinct (1992)
Fugitive, The (1993)
--------------------------------------------
Bottom 3 films:
-----------------
Ghost and the Darkness, The (1996)
Feeling Minnesota (1996)
Scream (1996)
*****************************************************
root@ima-app:/var/www/Hussam/A7#

\end{lstlisting}

I mostly identify with user 182. Moving forward I will call 182 ``my substitute''.


2. In order to find which 5 users are most correlated to the my substitute, 182, I utilized the code in the file ``recommendations.py'' from PCI book, Programming Collective Intelligence. I should have used the code for part one also, but I was trying to do everything from scratch. I modified the function ``topMatches'' and used it to find the 5 users that are most and least correlated to the my substitute, 182. I simply added an argument ``m'' to the function. if m = 1 or True, the function is not changed at all. If m = 0 or False is passed to the function, the function does not reverse the list after sorting which gives the least 5 correlated users. Below is the function ``topMatches'' after it has been modified.

\begin{lstlisting}[language=python, breakatwhitespace=〈false), label=topMatches, caption=topMatches]
def topMatches(m, prefs, person, n=5, similarity=sim_pearson):
    scores = [(similarity(prefs, person, other), other) for other in prefs
              if other != person]
    scores.sort()
    if m:
        scores.reverse()
    return scores[0:n]
\end{lstlisting}

I wrote a python script ``findcorrelated.py'' that takes a command line argument, user\_id, as input and prints out the 5 users that are most and least correlated to the user\_id passed to the script.

\lstinputlisting[language=Python, breakatwhitespace=〈false), label=The content of findcorrelated.py, caption= The content of findcorrelated.py]{Q1/findcorrelated.py}

After running the script ``findcorrelated.py'' and passing my substitute's id 182, I got the following users:

\begin{lstlisting}[language=bash, breakatwhitespace=〈false), label=Running findcorrelated.py, caption=Running findcorrelated.py]
root@ima-app:/var/www/Hussam/A7# python findcorrelated.py 182
Five Most Correlated Users to My Substitute: 182
--------------------------------------------------------
283
156
777
603
826
++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Five Least Correlated Users to My Substitute: 182
--------------------------------------------------------
127
132
492
617
641
root@ima-app:/var/www/Hussam/A7#
\end{lstlisting}


3. In order to compute ratings for all the films that the my substitute
has not seen and provide a list of the top/bottom 5 recommendations for films
that my substitute you should/shouldn't see. I used the function ``getRecommendations'' in ``recommendations.py'' to get top and bottom recommendations based on collaborative filtering.
 
I wrote a python script ``findrecommended55.py'' that takes user\_id as a command line argument, in this case my substitute user 182, and prints out the top and bottom 5 recommended films.

\lstinputlisting[language=Python, breakatwhitespace=〈false), label=The content of findrecommended55.py, caption= The content of findrecommended55.py]{Q1/findrecommended55.py}

After running the script ``findrecommended55.py'' and passing my substitute's id 182, I got the following output which has both lists of top and bottom 5 recommendations for my substitute, user 182:

\begin{lstlisting}[language=bash, breakatwhitespace=〈false), label=Running findrecommended55.py, caption=Running findrecommended55.py]
root@ima-app:/var/www/Hussam/A7# python findrecommended55.py 182
Five Most recommended Movies for My Substitute: 182
--------------------------------------------------------
They Made Me a Criminal (1939)
The Deadly Cure (1996)
Someone Else's America (1995)
Saint of Fort Washington, The (1993)
Prefontaine (1997)
++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Five Least Recommended Movies for My Substitute: 182
--------------------------------------------------------
Amityville II: The Possession (1982)
Amityville Curse, The (1990)
Amityville 3-D (1983)
Amityville 1992: It's About Time (1992)
3 Ninjas: High Noon At Mega Mountain (1998)
++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
root@ima-app:/var/www/Hussam/A7#
\end{lstlisting}


4. It was easy to find my favorite movie from the data, but so hard to find the least favorite movie because if I decide to watch a movie, I give it 5 minutes to get my attention. If it did not, I look for something else and forget the name, actors, ...etc and completely remove it from my memory. Anyways, I looked up one of the movies and read its story and it seems so boring, so I decided to choose it as my least favorite movie without seeing it. 

In order to generate a list of the top 5 most correlated and bottom 5 least correlated films to the two films, I have chosen.

My absolute favorite film of all times is: Braveheart (1995)

My least favorite film is: Amityville II: The Possession (1982)

I wrote a python script, findmyrecommendations.py. It imports ``recommendations.py'' and uses the function ``calculateSimilarItems'' with a slight modification to compute a list of recommended/not-recommended movies for me to watch/not-watch. The modification is done by passing an extra argument in the function, m. This argument is, then, passed to the function ``topMatches'' called from ``calculateSimilarItems''. If m = 1 or True, the function returns a list of recommended movies to watch. Similarly, if m = 0 or False, the function returns a list of non-recommended movies to not watch. Below is the function ``calculateSimilarItems'' after it has been modified.

\begin{lstlisting}[language=python, breakatwhitespace=〈false), label=topMatches, caption=topMatches]
def calculateSimilarItems(m, prefs, n=10):
    result = {}
    # Invert the preference matrix to be item-centric
    itemPrefs = transformPrefs(prefs)
    c = 0
    for item in itemPrefs:
        # Status updates for large datasets
        c += 1
        if c % 100 == 0:
            print '%d / %d' % (c, len(itemPrefs))
        # Find the most similar items to this one
        scores = topMatches(m, itemPrefs, item, n=n, similarity=sim_distance)
        result[item] = scores
    return result
\end{lstlisting}

Below is the code in ``findmyrecommendations.py''.

\lstinputlisting[language=Python, breakatwhitespace=〈false), label=The content of findmyrecommendations.py, caption= The content of findmyrecommendations.py]{Q1/findmyrecommendations.py}

After running the script ``findmyrecommendations.py'', I got the list of top and bottom 5 recommendations for myself:

\begin{lstlisting}[language=bash, breakatwhitespace=〈false), label=Running findmyrecommendations.py, caption=Running findmyrecommendations.py]
root@ima-app:/var/www/Hussam/A7# python findmyrecommendations.py
100 / 1664
200 / 1664
300 / 1664
400 / 1664
500 / 1664
600 / 1664
700 / 1664
800 / 1664
900 / 1664
1000 / 1664
1100 / 1664
1200 / 1664
1300 / 1664
1400 / 1664
1500 / 1664
1600 / 1664
100 / 1664
200 / 1664
300 / 1664
400 / 1664
500 / 1664
600 / 1664
700 / 1664
800 / 1664
900 / 1664
1000 / 1664
1100 / 1664
1200 / 1664
1300 / 1664
1400 / 1664
1500 / 1664
1600 / 1664
Best Recommended movies to watch based on Braveheart (1995):
---------------------------------------------------------------
Wings of Courage (1995)
Wife, The (1995)
The Deadly Cure (1996)
Sexual Life of the Belgians, The (1994)
Of Love and Shadows (1994)
++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Least Recommended movies to watch based on Braveheart (1995):
---------------------------------------------------------------
Aiqing wansui (1994)
American Strays (1996)
August (1996)
B. Monkey (1998)
Ballad of Narayama, The (Narayama Bushiko) (1958)
++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Best Recommended movies to watch based on Amityville II: The Possession (1982):
---------------------------------------------------------------
Yankee Zulu (1994)
Women, The (1939)
Woman in Question, The (1950)
Withnail and I (1987)
Winter Guest, The (1997)
++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Least Recommended movies to watch based on Amityville II: The Possession (1982):
---------------------------------------------------------------
'Til There Was You (1997)
8 Heads in a Duffel Bag (1997)
A Chef in Love (1996)
Above the Rim (1994)
Adventures of Pinocchio, The (1996)
++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
root@ima-app:/var/www/Hussam/A7#
\end{lstlisting}

I honestly watch less than 5 movies a year. I have not watched the majority of these movies, but after reading about them, I mostly agree with both lists of least recommended movies, and mostly disagree with movies in the list of recommended movies. 

\subsection*{Included Files:}
closest3users.txt, findclosestusers.py, findcorrelated.py, findmyrecommendations.py, findrecommended55.py, findtopbottom343.py, recommendations.py, recommendations.pyc, topbottom343.txt, u.data, u.item, u.user

Folder: data has u.data, u.item, u.user, and a sub-folder: movielens that has u.data, u.item, u.user.

\end{document}
